The Dynamics of Normalization Reweighting (target journal: Neuron)

### 2022-06-14 manuscript.Rmd
Added AQ scoring

processdata <- 0  # this flag determines the amount of processing, with 4 levels:
# 0 - do no processing, generate the pdf using existing versions of all scripts
# 1 - generate figures using the processed group data (requires XGB of storage)
# 2 - download processed individual data and analyse (requires XGB of storage)
# 3 - download all raw data and analyse (requires XGB of storage)

@@

knitr::opts_chunk$set(echo = TRUE)
colpal <- c('#000000','#8783CF','#FE5000','#004731','#603FB3')

@@

plotlims <- c(1,20,0,3)  # define the x and y limits of the plot (minx,maxx,miny,maxy)
ticklocsx <- seq(2,20,2)    # locations of tick marks on x axis
ticklocsy <- c(0,0.4,0.8,1,2,3)    # locations of tick marks on y axis
ticklabelsx <- seq(2,20,2)        # set labels for x ticks
ticklabelsy <- c(0,0.2,0.4,'','','')    # set labels for y ticks
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Frequency (Hz)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude (µV)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

@@

 }
if (processdata>1){    
d <- dir(path = 'temp/EEGprocessed', pattern = '*.RData', full.names=TRUE)
EEGtimecourse5Hz <- array(0,dim=c(length(d),4,8000))
EEGtimecourse7Hz <- array(0,dim=c(length(d),4,8000))
allwaves <- array(0,dim=c(length(d),4,9000))
allspec <- array(0,dim=c(length(d),4,duration*20))
allheads5 <- array(0,dim=c(length(d),4,64))
allheads7 <- array(0,dim=c(length(d),4,64))
for (s in 1:length(d)){
  load(d[s])
  targetchannels <- match(targetelectrodes, electrodes) - 2
  for (cond in 1:4){
    temp <- rowMeans(subjwaves[cond,,targetchannels],na.rm=TRUE)
    nainds <- which(is.na(temp))   # find occasional NaN values in the waveform
    # temp[nainds] <- (temp[nainds-1] + temp[nainds+1])/2   # interpolate adjacent values
    temp[nainds] <- 0   # replace with zeros
    allwaves[s,cond,] <- temp - mean(temp)
    spec <- fft(temp[1501:7500])/6000
    allspec[s,cond,] <- spec[1+(1:(duration*20))]

@@

@@

save(file='temp/EEGsummary.RData',list=c('allheads5','allheads7','allspec','allwaves','EEGtimecourse5Hz','EEGtimecourse7Hz','electrodes'))
 }

@@

meanspec[cond,] <- abs(colMeans(allspec[which(isincluded[cond,]==1),cond,],na.rm=TRUE))
meanwave[cond,] <- colMeans(allwaves[which(isincluded[cond,]==1),cond,],na.rm=TRUE)
meanhead5[cond,] <- abs(colMeans(allheads5[which(isincluded[cond,]==1),cond,],na.rm=TRUE))
meanhead7[cond,] <- abs(colMeans(allheads7[which(isincluded[cond,]==1),cond,],na.rm=TRUE))
}

@@

polygon(c(1/6,freqlist,20),c(0,2*meanspec[1,],0)+3,border=NA,col=colpal[1])
polygon(c(1/6,freqlist,20),c(0,2*meanspec[2,],0)+2,border=NA,col=colpal[1])
polygon(c(1/6,freqlist,20),c(0,2*meanspec[3,],0)+1,border=NA,col=colpal[2])
polygon(c(1/6,freqlist,20),c(0,2*meanspec[4,],0),border=NA,col=colpal[3])

lines(freqlist,2*meanspec[1,]+3,lwd=2,col='black')
lines(freqlist,2*meanspec[2,]+2,lwd=2,col='black')
lines(freqlist,2*meanspec[3,]+1,lwd=2,col='black')
lines(freqlist,2*meanspec[4,],lwd=2,col='black')

@@

  text(-0.02,0.61,'5Hz',pos=4,cex=2)
  text(-0.02,0.89,'7Hz',pos=4,cex=2)

>>>>

processdata <- 1  # this flag determines the amount of processing, with 4 levels:
# 0 - do no processing, generate the pdf using existing versions of all scripts
# 1 - generate figures using the processed group data (requires XGB of storage)
# 2 - download processed individual data and analyse (requires XGB of storage)
# 3 - download all raw data and analyse (requires XGB of storage)

@@

scoreAQ <- function(qdata){
# score the AQ questionnaire
# input is a row from a spreadsheet indicating responses to 50 questions
AQanswers <- toupper(c('(1) Definitely agree','(2) Slightly agree', '(3) Slightly disagree', '(4) Definitely disagree'))
reverseitems <- rep(0,50)
 # reverse scored questionnaire items
reverseitems[c(3,8,10,11,14,15,17,24,25,27,28,29,30,31,32,34,36,37,38,40,44,47,48,49,50)] <- 1

  thisAQ <- 0
  nansweredAQ <- 0
  for (q in 1:50){
    response <- which(AQanswers==toupper(as.character(qdata[1,q])))
    if (length(response)>0){
      nansweredAQ <- nansweredAQ + 1
      if (reverseitems[q]==0){response <- 5 - response}
      if (response>2){thisAQ <- thisAQ + 1}
    }}
thisAQ <- round(50*thisAQ/nansweredAQ)    
return(thisAQ)}

knitr::opts_chunk$set(echo = TRUE)
colpal <- c('#000000','#8783CF','#FE5000','#228B22','#8B8000')

@@

plotlims <- c(0,20,0,3)  # define the x and y limits of the plot (minx,maxx,miny,maxy)
ticklocsx <- seq(0,20,5)    # locations of tick marks on x axis
ticklocsy <- c(0,0.4,0.8,1,2,3)    # locations of tick marks on y axis
ticklabelsx <- seq(0,20,5)        # set labels for x ticks
ticklabelsy <- c(0,0.2,0.4,'','','')    # set labels for y ticks
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Frequency (Hz)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude (µV)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

@@

 }
if (processdata>1){    
d <- dir(path = 'temp/EEGprocessed', pattern = '*.RData', full.names=TRUE)
EEGtimecourse5Hz <- array(0,dim=c(length(d),4,8000))
EEGtimecourse7Hz <- array(0,dim=c(length(d),4,8000))
allwaves <- array(0,dim=c(length(d),4,9000))
allspec <- array(0,dim=c(length(d),4,duration*20))
allheads5 <- array(0,dim=c(length(d),4,64))
allheads7 <- array(0,dim=c(length(d),4,64))
allAQ <- NULL
for (s in 1:length(d)){
  load(d[s])
  targetchannels <- match(targetelectrodes, electrodes) - 2
  for (cond in 1:4){
    temp <- rowMeans(subjwaves[cond,,targetchannels],na.rm=TRUE)
    nainds <- which(is.na(temp))   # find occasional NaN values in the waveform
    # temp[nainds] <- (temp[nainds-1] + temp[nainds+1])/2   # interpolate adjacent values
    temp[nainds] <- 0   # replace with zeros
    allwaves[s,cond,] <- temp - mean(temp)
    spec <- fft(temp[1501:7500])/6000
    allspec[s,cond,] <- spec[1+(1:(duration*20))]

@@

allAQ[s] <- scoreAQ(qdatasubj[1,5:54])

@@

save(file='temp/EEGsummary.RData',list=c('allheads5','allheads7','allspec','allwaves','EEGtimecourse5Hz','EEGtimecourse7Hz','electrodes','allAQ'))
 }

@@

meanspec[cond,] <- colMeans(abs(allspec[which(isincluded[cond,]==1),cond,]),na.rm=TRUE)
meanwave[cond,] <- colMeans(allwaves[which(isincluded[cond,]==1),cond,],na.rm=TRUE)
meanhead5[cond,] <- abs(colMeans(abs(allheads5[which(isincluded[cond,]==1),cond,]),na.rm=TRUE))
meanhead7[cond,] <- abs(colMeans(abs(allheads7[which(isincluded[cond,]==1),cond,]),na.rm=TRUE))
}

@@

polygon(c(1,freqlist[6:120],20),c(0,2*meanspec[1,6:120],0)+3,border=NA,col=colpal[1])
polygon(c(1,freqlist[6:120],20),c(0,2*meanspec[2,6:120],0)+2,border=NA,col=colpal[1])
polygon(c(1,freqlist[6:120],20),c(0,2*meanspec[3,6:120],0)+1,border=NA,col=colpal[2])
polygon(c(1,freqlist[6:120],20),c(0,2*meanspec[4,6:120],0),border=NA,col=colpal[3])

lines(freqlist[6:120],2*meanspec[1,6:120]+3,lwd=2,col='black')
lines(freqlist[6:120],2*meanspec[2,6:120]+2,lwd=2,col='black')
lines(freqlist[6:120],2*meanspec[3,6:120]+1,lwd=2,col='black')
lines(freqlist[6:120],2*meanspec[4,6:120],lwd=2,col='black')

@@

  text(-0.02,0.89,'5Hz',pos=4,cex=2)
  text(-0.02,0.61,'7Hz',pos=4,cex=2)
